---
title: "Seahtrue"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
```{r, echo=FALSE}

path <- here::here("inst", "extdata", "seahtrue_banner.png")
htmltools::img(src = knitr::image_uri(path), 
               alt = 'logo',
               style = 'width: 800px; height: 600x') # Added width and height as CSS
          
```

```{r, include = FALSE}
# Change the default values of chunk options in a document.
# Include = FALSE prevents code and results from appearing in the finished file. 
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r setup, include=FALSE}
# The option echo=TRUE causes your code to be included in the knitted document. 
knitr::opts_chunk$set(echo = TRUE)
```

# General information.

   * This page serves as a quick guide for our package Seahtrue.

   * For extra information about the Seahtrue package you can visit our package documentation.
   
## Overview

Key sections covered in this guide include:

#### Installation
Instructions for installing Seahtrue.

#### Usage
An introduction to the core function, run_seahtrue.

#### Seahtrue Workflow (Version 0.1.0)
An overview of how Seahtrue works including loading, preprocessing and validating provided seahorse data.

#### Preprocess Output
Details about the data generated after preprocessing.

#### Examples
Practical examples of Seahtrue in action.

## Installation

You can install the latest version of Seahtrue with the following command:

``` {r install_github}
devtools::install_github("vcjdeboer/seahtrue", ref = "develop-gerwin")
```

When asking if you would like to update packages, select 1.
This will make sure our package works with the installed packages on the system.
Make sure that these updated packages will not interfere with any current R projects you have.

After installation of our seahtrue package and updating all packages to the required versions, 
you will have to restart R. This to make sure all the required package versions will be in 
use. Next our package will be loaded.

``` {r load_seahtrue_package}
library(seahtrue)
```


## Usage
Version 0.1.0 of the Seahtrue package contains a run_seahtrue function, which is
located at the master.R in the /R directory. This function performs all the required
steps for the current version of this package.

### Step 1 and 2: Loading and validating input.

This function loads a Seahorse Excel 
file as an input, which requires the full path to the Excel file. The Excel input 
will be validated. This to make sure it contains Seahorse data required for our 
analysis.

### Step 3 and 4: Preprocessing and validating preprocessed data.

The next step of our analysis is to preprocess the Seahorse Excel data that is 
loaded and validated during the previous step. This preprocess step produces a dataset 
with preprocessed data. 

The preprocessed dataset is then validated to check if the seahorse data meets
our criteria. When Seahorse data contains unexpected data, which doesn't meet our 
validation criteria, a warning will be given. 

Note that the Seahtrue analysis will still be performed with validation that 
doesn't meet our criteria.
In summary preprocessing is a mandatory step to make sure our Seahtrue calculation 
will work well.

## Seahtrue workflow
The main workflow of seahtrue will run when using the function run_seahtrue.
This function currently fulfills 4 different purposes:

1. Validates the input data                                   
2. Reads the input data                                  
3. Preprocesses the input data
4. Validates the preprocess output

## Preprocess output
The eventual output is a preprocessed dataset.

| Preprocessing steps                                          | 
|--------------------------------------------------------------|
| changing columns names                                       |
| adding new time columns                                      |
| adding injection info columns                                |
| add plate_id column                                          |
| calculating background data                                  |
| calculating raw pH emission data                             |

The following output is returned by step 3. 
   
|      Objects       |                                         Description                                            |
|--------------------|------------------------------------------------------------------------------------------------|
| plate_id           | Barcode from the 'Assay Configuration' sheet from the Seahorse Excel file.                     |
| filepath_seahorse  | Path to Seahorse Excel file.                                                                   |
| date               | Date derived from 'Assay Configuration' sheet of the Seahorse Excel file.                      |
| assay_info         | Dataframe with information from the 'Assay Configuration' sheet from the Seahorse Excel file.  |
| raw_info           | Dataframe with information from the 'Raw' sheet from the Seahorse Excel file.                  |
| injection_info     | Info from Operation log sheet.                                                                 |
| rate_data          | With information from the 'Rate' sheet from the Seahorse Excel file.                           |

## Examples

Here are some examples of the preprocessed output data of version 0.1.0 of our 
Seahtrue package. We use PBMC seahorse Excel data as an input. These Excel data were 
converted using the Seahorse .asyr output generated by the Agilent Wave Software.

To make sure our run_seahtrue function runs properly we can use the following 
PBMC datasets:

| * 20191219_SciRep_PBMCs_donor_A.xlsx                                    
| * 20200110_SciRep_PBMCs_donor_B.xlsx                                    
| * 20200110_SciRep_PBMCs_donor_C.xlsx                                    

These seahorse datasets are located at the inst/extdata directory. 
When installing seahtrue these dataset will be located at the installation 
location of seahtrue, and to access this location we have to make use of 
system.file() provided by 'base' R.

### Example 1: Run PBMC dataset donor A

``` {r run_seahtrue_test_donor_A}
donor_A <- seahtrue::run_seahtrue(filepath_seahorse = 
                         system.file("extdata", "20191219_SciRep_PBMCs_donor_A.xlsx", package = "seahtrue"))
donor_A
```

#### Plate ID
Barcode from the 'Assay Configuration' sheet from the Seahorse Excel file.
``` {r plate_id}
donor_A$plate_id
```

#### Filepath
Path to Seahorse Excel file
``` {r filepath_seahorse}
donor_A$filepath_seahorse
``` 

#### Date
Date derived from 'Assay Configuration' sheet of the Seahorse Excel file.
``` {r date}
donor_A$date
```
#### Assay information
Dataframe with information from the 'Assay Configuration' sheet from the Seahorse Excel file.
``` {r assay_info}
donor_A$assay_info
```

#### Injection information
Dataframe with information derived from 'Operation log' sheet.
``` {r injection_info}
donor_A$injection_info
```

#### Preprocessed Raw data
Preprocessed raw dataframe from Seahorse 'Raw' sheet
``` {r raw_data}
donor_A$raw_data
```

#### Preprocessed Rate data
Preprocessed rate data from 'Rate' sheet
``` {r rate_data}
donor_A$rate_data
```

### Example 2: PBMC donor A output
``` {r seahorse_output_donor_A}
donor_A <- seahtrue::seahtrue_output_donor_A
donor_A
```

### Example 3: PBMC donor A output

``` {r seahorse_output_donor_B}
donor_B <- seahtrue::seahtrue_output_donor_B
donor_B
```

### Example 4: PBMC donor C output
``` {r seahorse_output_donor_C}
donor_C <- seahtrue::seahtrue_output_donor_C

donor_C
```

## Upcoming
We can use the validated preprocess output data for our calculations. 
Calculations will be released during the next official release of Seahtrue.
The output of our calculations will be visualized. For visualization we can use Shiny.

## Getting Help
If you need assistance or want to learn more about the Seahtrue package you can
utilize the documentation page of our package (?seahtrue). This will provide 
you with links to the documentation of our different functions as well as links
to some general Seahorse background information related to our package.

## Contributing and Feedback
Suggestions and improvements are welcome! If you have any feedback on our Seahtrue 
package when it comes to improvements or bugs, let us know.
Your input is valuable and can help us make Seahtrue better.
