---
title: "Seahtrue"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
 
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
```

seahtrue is a package for analyzing extracellular flux data. It organizes data from excel files into nested tibbles and allows for multi-experiment analysis detailed raw data processing.

### Installation

You can install the latest version of seahtrue with the following command:

```{r install_github, message=TRUE, warning=FALSE}
devtools::install_github("vcjdeboer/seahtrue", ref = "main", quiet = TRUE)
```

### Getting Help
If you need assistance or want to learn more about the Seahtrue package you can
utilize the documentation page of our package (?seahtrue). This will provide 
you with links to the documentation of our different functions as well as links
to some general Seahorse background information related to our package.

### Contributing and Feedback
Suggestions and improvements are welcome! If you have any feedback on our Seahtrue 
package when it comes to improvements or bugs, let us know.
Your input is valuable and can help us make Seahtrue better.

