---
output: rmarkdown::html_vignette
description: |
  Overview of the Seahtrue package.
vignette: >
  %\VignetteIndexEntry{Introduction to Seahtrue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
---
## The seahtrue package

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  rmarkdown.html_vignette.check_title = FALSE
)
```

### Seahtrue overview

The seahtrue package offers a set of functions to be able to perform reproducible data analysis of extracellular flux analysis. The main function `revieve_xfplate()` reads, preprocess and validates the input data and outputs experimental details and outcome variables in an organized (tidy) way. The output of the `revive_xfplate()` is a nested tibble.

### Extracellular flux analysis scientific primer

With instruments such as the Seahorse XF analyzer from Agilent, but also the O2K from Oroboros, other oyxgraphs (from Hansatech instruments for example) and the ReSipher from Lucid Scientific, scientists are able to analyze oxygen consumption of living biological samples. 

Oxygen consumption of cells or small model organism can provide insights into the function of the mitochondria, since mitochondria are most of the time the main O2 consumers of cells. Apart form oxygen consumption the Seahorse XF analyzer is able to analyze in parallel the extracellular acidification of the culture medium in which the sample is emerged. This can be a proxy for glycolytic activity of samples. 

Seahorse extracellular flux instruments performs analysis of O2 and pH in either 96 wells, 24 wells or 8 wells, and typically O2 and pH are monitored over a period of around 1 hour, in discrete measurements of typically 3 minutes each. Furthermore, perturbations of cellular functional states can be induced by adding compounds while performing the assay. The most common perturbations that are performed are injections of oligomycin, fccp and anitmycin/rotenone, known as a mitostress test. 

### Getting started

The `tidyverse` and `seahtrue` package are used here, so we have to load them first. Also, the `here` package can be usefull.

```{r load packages}
#suppressPackageStartupMessages({
  library(tidyverse)
  library(seahtrue)
  library(here)
#})

# in your own workflow you don't have to supresses messages 
# off course, this is used here only for cleaner vignettes


```

The `revive_xfplate()` takes the .xlsx output file from a 96 well XF experiment as its input.

```{r read_plate_data, eval=FALSE, include=TRUE}
my_file_path <- "/Users/myname/data/my_xf_plate.xlsx" # mac style folder path

my_plate <- my_file_path %>% revive_xfplate()

```

Handling files in your preferred IDE, such as RStudio, can be a challenge. Working within your own Rstudio project is often advisable, where you can put your xlsx files into a `data` folder. If this is your workflow you can use relative paths, for example like this:

```{r file handling, eval=FALSE, include=TRUE}
my_file_path <- here::here("data", "my_xf_plate. xlsx")

my_plate <- my_file_path %>% revive_xfplate()

```

Otherwise, when you are on a Mac you might want to copy the direct path from the finder path bar, by right clicking on the file and `copy my_xf_plate.xlsx as pathname`. You need to have the finder path bar set to `Show` for this. 

In this vignette we use an xlsx file that we included with the seahtrue package:

```{r example revive}
my_file_path <- system.file("extdata", "20191219_SciRep_PBMCs_donor_A.xlsx", package = "seahtrue")

my_plate <- my_file_path %>% revive_xfplate()

```


Information received from the 'Raw' Seahorse datasheet.

|     Constant     |  Description |
|------------------|--------------|
| Measurement      |              |
| Tick             |              |
| Well             |              |
| Group            |              |
| TimeStamp        |              |
| Well Temperature |              |
| Env. Temperature |              |
| O2 is Valid      |              |
| O2 (MmHg)        |              |
| O2 Light Emision |              |
| O2 Dark Emission |              |
| O2 Ref Light     |              |
| O2 Corrected Em  |              |
| pH is Valid, pH  |              |
| pH Light         |              |
| pH Dark          |              |
| pH Ref Light     |              |
| pH Ref Dar       |              |
| pH Corrected Em. |              |

Information received from the 'Assay information' Seahorse datasheet.

| Constant | Description |
|----------|-------------|
| F0       |             |
| V_C      |             |
| Tau_AC   |             |
| Tau_W    |             |
| Tau_C    |             |
| Tau_P    |             |
| KSV      |             |
| gain1    |             |
| gain2    |             |
| pH_0     |             |

